---
title: "5543 project"
format: html
execute:
  echo: true
  warning: false
  error: true
  message: false
---

```{r}
# Setup chunk: install packages if needed (commented by default)
# install.packages(c("tidyverse","survey","srvyr","scales","glue"))

library(tidyverse)
library(survey)
library(srvyr)
library(scales)
library(glue)
library(plotly)
library(patchwork)
library(nnet)
library(dplyr)
library(tidyr)
library(car)
library(brglm2)
library(logistf)
library(visreg)
library(broom)
```

```{r}
#load data
household <- read_csv("data/CSV/SIH19BH.csv", show_col_types = FALSE)
income <- read_csv("data/CSV/SIH19BI.csv", show_col_types = FALSE)
loans <- read_csv("data/CSV/SIH19BL.csv", show_col_types = FALSE)
person <- read_csv("data/CSV/SIH19BP.csv", show_col_types = FALSE)

#load older data
household2013 <- read_csv("data/CSV/SIH13BH.csv", show_col_types = FALSE)
income2013 <- read_csv("data/CSV/SIH13BI.csv", show_col_types = FALSE)
loans2013 <- read_csv("data/CSV/SIH13BL.csv", show_col_types = FALSE)
person2013 <- read_csv("data/CSV/SIH13BP.csv", show_col_types = FALSE)
```


```{r}
#Aggregate loans (household-level debt measures)

loans_agg <- loans %>%
  group_by(ABSHID) %>%
  summarise(
    hh_total_weekly_repayments = sum(WKREPLN, na.rm = TRUE),
    hh_total_outstanding = sum(AMTOLN, na.rm = TRUE),
    hh_num_loans = n(),
    .groups = "drop"
  )


loans_agg2013 <- loans2013 %>%
  group_by(ABSHID) %>%
  summarise(
    hh_total_weekly_repayments = sum(WKREPLN, na.rm = TRUE),
    hh_total_outstanding = sum(AMTOLN, na.rm = TRUE),
    hh_num_loans = n(),
    .groups = "drop"
  )
```

```{r}
#Aggregate IU incomes to household (optional check)

iu_agg <- income %>%
  group_by(ABSHID) %>%
  summarise(
    hh_income_from_ius = sum(INCTSCU8, na.rm = TRUE),
    hh_num_ius = n(),
    .groups = "drop"
  )

iu_agg2013 <- income2013 %>%
  group_by(ABSHID) %>%
  summarise(
    hh_income_from_ius = sum(INCTSCU8, na.rm = TRUE),
    hh_num_ius = n(),
    .groups = "drop"
  )


```


```{r}
#Extract household-reference-person health (to attach a person-level wellbeing proxy)

person_head <- person %>%
  filter(HHPOS == 1) %>%     # household reference person
  select(ABSHID, SAHQ01, AGEEC, SEXP, LVLEDUA) %>%
  mutate(
    head_poor_health_flag = case_when(
      SAHQ01 %in% c(4,5) ~ 1L,  # 4 = Fair, 5 = Poor
      TRUE ~ 0L
    )
  )
person_head2013 <- person2013 %>%
  filter(HHPOS == 1) %>%     # household reference person
  select(ABSHID, DISBPER, AGEBC, SEXP) %>%
  mutate(
    head_poor_health_flag = case_when(
      DISBPER == 2 ~ 1L,
      TRUE ~ 0L
    )
  )

```

```{r}
#merge tables into a single household dataset

hh_merged <- household %>%
  left_join(loans_agg, by = "ABSHID") %>%
  left_join(iu_agg, by = "ABSHID") %>%
  left_join(person_head, by = "ABSHID")


hh_merged2013 <- household2013 %>%
  left_join(loans_agg2013, by = "ABSHID") %>%
  left_join(iu_agg2013, by = "ABSHID") %>%
  left_join(person_head2013, by = "ABSHID")

#rename variables
rename_map <- c(
  "SIHHHWT"   = "hh_weight",
  "HCOSTH8A"  = "weekly_housing_cost_pct",
  "HCOST10A"  = "weekly_housing_cost_cat",
  "INCTSCH8"  = "hh_total_weekly_income",
  "DISPSCH8"  = "hh_weekly_disposable_income",
  "LIFECYCH"  = "life_cycle_group_code",
  "DPKIDHBC"  = "num_dependents_under25",
  "FSTRSS19"  = "num_financial_stress_events",
  "EMGMONEY"  = "ability_raise_2000_flag"
)

for(orig in names(rename_map)){
  if(orig %in% names(hh_merged)){
    hh_merged <- hh_merged %>% rename(!!rename_map[[orig]] := all_of(orig))
  }
}

for(orig in names(rename_map)){
  if(orig %in% names(hh_merged2013)){
    hh_merged2013 <- hh_merged2013 %>% rename(!!rename_map[[orig]] := all_of(orig))
  }
}
```




```{r}
#Recode life-cycle group to readable labels & compute derived vars

hh_merged <- hh_merged %>%
  mutate(
    life_cycle_group = case_when(
      life_cycle_group_code == 0  ~ "Not applicable",
      life_cycle_group_code == 1  ~ "Lone person <35",
      life_cycle_group_code == 2  ~ "Couple (ref <35)",
      life_cycle_group_code == 3  ~ "Couple w dependents, eldest <5",
      life_cycle_group_code == 4  ~ "Couple w dependents, eldest 5-14",
      life_cycle_group_code == 5  ~ "Couple w dependents, eldest 15-24",
      life_cycle_group_code == 6  ~ "One parent w dependents",
      life_cycle_group_code == 7  ~ "Couple w dep & non-dep",
      life_cycle_group_code == 8  ~ "Couple w non-dep only",
      life_cycle_group_code == 9  ~ "Couple (ref 55-64)",
      life_cycle_group_code == 10 ~ "Couple (ref 65+)",
      life_cycle_group_code == 11 ~ "Lone person 65+",
      TRUE ~ paste0("Other (", life_cycle_group_code, ")")
    ),
    housing_stress_flag = case_when(
      !is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct > 30 ~ 1L,
      !is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct <= 30 ~ 0L,
      TRUE ~ NA_integer_
    ),
    housing_stress_flag_20 = case_when(
      !is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct > 20 ~ 1L,
      !is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct <= 20 ~ 0L,
      TRUE ~ NA_integer_
    ),
    housing_stress_flag_40 = case_when(
      !is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct > 40 ~ 1L,
      !is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct <= 40 ~ 0L,
      TRUE ~ NA_integer_
    ),
    housing_stress_flag_50 = case_when(
      !is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct > 50 ~ 1L,
      !is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct <= 50 ~ 0L,
      TRUE ~ NA_integer_
    ),
    ability_raise_2000 = case_when(
      ability_raise_2000_flag == 1 ~ "Can raise $2000",
      ability_raise_2000_flag == 5 ~ "Cannot raise $2000",
      TRUE ~ NA_character_
    )
  )

hh_merged2013 <- hh_merged2013 %>%
  mutate(
    life_cycle_group = case_when(
      life_cycle_group_code == 0  ~ "Not applicable",
      life_cycle_group_code == 1  ~ "Lone person <35",
      life_cycle_group_code == 2  ~ "Couple (ref <35)",
      life_cycle_group_code == 3  ~ "Couple w dependents, eldest <5",
      life_cycle_group_code == 4  ~ "Couple w dependents, eldest 5-14",
      life_cycle_group_code == 5  ~ "Couple w dependents, eldest 15-24",
      life_cycle_group_code == 6  ~ "One parent w dependents",
      life_cycle_group_code == 7  ~ "Couple w dep & non-dep",
      life_cycle_group_code == 8  ~ "Couple w non-dep only",
      life_cycle_group_code == 9  ~ "Couple (ref 55-64)",
      life_cycle_group_code == 10 ~ "Couple (ref 65+)",
      life_cycle_group_code == 11 ~ "Lone person 65+",
      TRUE ~ paste0("Other (", life_cycle_group_code, ")")
    ))

```


```{r}
#Set up survey replicate design (household)
rep_cols <- grep("^WHS\\d{2}$", names(hh_merged), value = TRUE)

# create srvyr survey object using Fay method (rho = 0.5)
hh_svy <- hh_merged %>%
  as_survey_rep(weights = hh_weight,
                repweights = all_of(rep_cols),
                type = "Fay",
                rho = 0.5)

# Set up survey replicate design (household) - SIH 2013
rep_cols2013 <- grep("^WHS\\d{4}$", names(hh_merged2013), value = TRUE)

# Create srvyr survey object using Fay method (rho = 0.5)

hh_svy2013 <- hh_merged2013 %>%
  as_survey_rep(
    weights = hh_weight,          
    repweights = all_of(rep_cols2013),
    type = "Fay",
    rho = 0.5
  )

```


```{r}
# Proportion of households in housing stress (>30%) by life-cycle group
stress_by_life <- hh_svy %>%
  group_by(life_cycle_group) %>%
  summarize(
    pct_stress = survey_mean(housing_stress_flag, vartype = c("ci","se"), na.rm = TRUE)
  ) %>%
  arrange(desc(pct_stress))

stress_by_life %>% collect() %>% print(n = Inf)


# Proportion of households in housing stress (>20%) by life-cycle group
stress_by_life_20 <- hh_svy %>%
  group_by(life_cycle_group) %>%
  summarize(
    pct_stress = survey_mean(housing_stress_flag_20, vartype = c("ci","se"), na.rm = TRUE)
  ) %>%
  arrange(desc(pct_stress))

stress_by_life_20 %>% collect() %>% print(n = Inf)


# Proportion of households in housing stress (>40%) by life-cycle group
stress_by_life_40 <- hh_svy %>%
  group_by(life_cycle_group) %>%
  summarize(
    pct_stress = survey_mean(housing_stress_flag_40, vartype = c("ci","se"), na.rm = TRUE)
  ) %>%
  arrange(desc(pct_stress))

stress_by_life_40 %>% collect() %>% print(n = Inf)

# Proportion of households in housing stress (>50%) by life-cycle group
stress_by_life_50 <- hh_svy %>%
  group_by(life_cycle_group) %>%
  summarize(
    pct_stress = survey_mean(housing_stress_flag_50, vartype = c("ci","se"), na.rm = TRUE)
  ) %>%
  arrange(desc(pct_stress))

stress_by_life_50 %>% collect() %>% print(n = Inf)
```


```{r}
plot1_df <- stress_by_life %>%
  collect() %>%
  mutate(
    pct = pct_stress * 100,
    low = pct_stress_low * 100,
    up  = pct_stress_upp * 100
  ) %>%
  arrange(pct) %>%
  mutate(life_cycle_group = factor(life_cycle_group, levels = unique(life_cycle_group)))

p1 <- ggplot(plot1_df, aes(x = life_cycle_group, y = pct)) +
  geom_col() +
  geom_errorbar(aes(ymin = low, ymax = up), width = 0.2) +
  coord_flip() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Proportion of Households in Housing Stress (>30% housing costs)",
    x = "Life-cycle group",
    y = "Percent of households (weighted, 95% CI)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 10)   
  )

# Save plot
ggsave("plot_housing_stress_by_lifecycle.png", p1, width = 10, height = 6, dpi = 300)


plot2_df <- stress_by_life_20 %>%
  collect() %>%
  mutate(
    pct = pct_stress * 100,
    low = pct_stress_low * 100,
    up  = pct_stress_upp * 100
  ) %>%
  arrange(pct) %>%
  mutate(life_cycle_group = factor(life_cycle_group, levels = unique(life_cycle_group)))

p2 <- ggplot(plot2_df, aes(x = life_cycle_group, y = pct)) +
  geom_col() +
  geom_errorbar(aes(ymin = low, ymax = up), width = 0.2) +
  coord_flip() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Proportion of Households in Housing Stress (>20% housing costs)",
    x = "Life-cycle group",
    y = "Percent of households (weighted, 95% CI)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 10)   
  )

plot3_df <- stress_by_life_40 %>%
  collect() %>%
  mutate(
    pct = pct_stress * 100,
    low = pct_stress_low * 100,
    up  = pct_stress_upp * 100
  ) %>%
  arrange(pct) %>%
  mutate(life_cycle_group = factor(life_cycle_group, levels = unique(life_cycle_group)))

p3 <- ggplot(plot3_df, aes(x = life_cycle_group, y = pct)) +
  geom_col() +
  geom_errorbar(aes(ymin = low, ymax = up), width = 0.2) +
  coord_flip() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Proportion of Households in Housing Stress (>40% housing costs)",
    x = "Life-cycle group",
    y = "Percent of households (weighted, 95% CI)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 10)   
  )


plot4_df <- stress_by_life_50 %>%
  collect() %>%
  mutate(
    pct = pct_stress * 100,
    low = pct_stress_low * 100,
    up  = pct_stress_upp * 100
  ) %>%
  arrange(pct) %>%
  mutate(life_cycle_group = factor(life_cycle_group, levels = unique(life_cycle_group)))

p4 <- ggplot(plot4_df, aes(x = life_cycle_group, y = pct)) +
  geom_col() +
  geom_errorbar(aes(ymin = low, ymax = up), width = 0.2) +
  coord_flip() +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  labs(
    title = "Proportion of Households in Housing Stress (>50% housing costs)",
    x = "Life-cycle group",
    y = "Percent of households (weighted, 95% CI)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 10)   
  )

# Print to the HTML output
ggplotly(p1)
ggplotly(p2)
ggplotly(p3)
ggplotly(p4)

```



```{r}
#Quick comparison: single-parent households vs others

single_parent_summary <- hh_svy %>%
  mutate(is_single_parent = (life_cycle_group == "One parent w dependents")) %>%
  group_by(is_single_parent) %>%
  summarize(
    pct_in_stress = survey_mean(housing_stress_flag, vartype = c("ci","se"), na.rm = TRUE),
    mean_hc_pct = survey_mean(weekly_housing_cost_pct, vartype = c("ci","se"), na.rm = TRUE)
  )

single_parent_summary %>% collect() %>% print()

```


```{r}
# --- State code mapping ---
state_labels <- c(
  "1" = "New South Wales",
  "2" = "Victoria",
  "3" = "Queensland",
  "4" = "South Australia",
  "5" = "Western Australia",
  "6" = "Tasmania",
  "7" = "Northern Territory",
  "8" = "Australian Capital Territory",
  "9" = "Other Territories"
)

purchase2013 <- hh_svy2013 %>%
  filter(!is.na(HVALPRIC), HVALPRIC > 0) %>%
  group_by(STATEHBC) %>%
  summarise(avg_price = survey_mean(HVALPRIC, na.rm = TRUE)) %>%
  mutate(
    year = 2013,
    STATE = state_labels[as.character(STATEHBC)]
  )
  


purchase2019 <- hh_svy %>%
  filter(!is.na(HVALPRIC), HVALPRIC > 0) %>%
  group_by(STATEHEC) %>%
  summarise(avg_price = survey_mean(HVALPRIC, na.rm = TRUE)) %>%
  mutate(
    year = 2019,
    STATE = state_labels[as.character(STATEHEC)]
  )

# --- Combine ---
purchase_compare <- bind_rows(purchase2013, purchase2019) 


p <- ggplot(purchase_compare, aes(x = year, y = avg_price/1000, group = STATE, color = STATE)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  # Note: I'm removing geom_text here as plotly handles tooltips better, 
  # but you can leave it if you want fixed labels.
  labs(
    title = "Trend in Average dwelling purchase prices by State (2013 to 2019)",
    x = "Year",
    y = "Average purchase price ($'000s)",
    color = "State"
  ) +
  scale_x_continuous(breaks = unique(purchase_compare$year)) +
  theme_minimal(base_size = 9) +
  theme(
    plot.title = element_text(size = 10, face = "bold")
  )

ggplotly(p)
```





```{r}
#Weekly housing costs (SIHC basis) - HH	HCOSTSH

# --- Compute weighted means only ---
cost_summary2013 <- hh_svy2013 %>%
  group_by(life_cycle_group) %>%
  summarise(
    mean_cost = survey_mean(HCOSTSH, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(year = 2013)

cost_summary2019 <- hh_svy %>%
  group_by(life_cycle_group) %>%
  summarise(
    mean_cost = survey_mean(HCOSTSH, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(year = 2019)

# --- Combine ---
cost_summary <- bind_rows(cost_summary2013, cost_summary2019)

# Ensure numeric (in case survey_mean makes it list-like)
cost_summary <- cost_summary %>%
  mutate(mean_cost = as.numeric(mean_cost)) %>% 
  filter(life_cycle_group != "Not applicable")



p <- ggplot(cost_summary, aes(x = life_cycle_group, y = mean_cost, fill = factor(year))) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  # Note: The fixed geom_text labels will still be present, but plotly's
  # tooltips provide a cleaner interactive experience.
  geom_text(aes(label = paste0("$", round(mean_cost))),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3) +
  labs(
    title = "Weekly housing costs across life-cycle groups (2013 vs 2019)",
    x = "Life-cycle group",
    y = "Mean weekly housing cost ($)",
    fill = "Year"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
    plot.title = element_text(size = 11, face = "bold")
  )

ggplotly(p)
```


```{r}
#Total current weekly IU income from all sources	INCTSCU8
# --- Weighted mean income 2013 ---
income_summary2013 <- hh_svy2013 %>%
  group_by(life_cycle_group) %>%
  summarise(
    mean_income = survey_mean(hh_income_from_ius, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(year = 2013)

# --- Weighted mean income 2019 ---
income_summary2019 <- hh_svy %>%
  group_by(life_cycle_group) %>%
  summarise(
    mean_income = survey_mean(hh_income_from_ius, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(year = 2019)

# --- Combine ---
income_summary <- bind_rows(income_summary2013, income_summary2019)

# Ensure numeric (survey_mean sometimes returns labelled structure)
income_summary <- income_summary %>%
  mutate(mean_income = as.numeric(mean_income)) %>% 
  filter(life_cycle_group != "Not applicable")


p <- ggplot(income_summary, aes(x = life_cycle_group, y = mean_income, fill = factor(year))) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = paste0("$", round(mean_income))),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 3) +
  labs(
    title = "Weekly household income across life-cycle groups (2013 vs 2019)",
    x = "Life-cycle group",
    y = "Mean weekly income ($)",
    fill = "Year"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 8),
    plot.title = element_text(size = 11, face = "bold")
  )

ggplotly(p)
```

```{r}
svy_afford_2013 <- hh_svy2013 %>%
  filter(!is.na(HCOSTSH), !is.na(hh_income_from_ius), hh_income_from_ius > 0) %>%
  mutate(afford_ratio = HCOSTSH / hh_income_from_ius)    # ratio in (0, Inf)

svy_afford_2019 <- hh_svy %>%
  filter(!is.na(HCOSTSH), !is.na(hh_income_from_ius), hh_income_from_ius > 0) %>%
  mutate(afford_ratio = HCOSTSH / hh_income_from_ius)

# ---- 2) Use survey-aware group summary to get mean ratio (estimate only) ----
# We use survey_mean(..., vartype = NULL) to request only the estimate column to avoid CI complications.
summary2013 <- svy_afford_2013 %>%
  group_by(life_cycle_group) %>%
  summarise(mean_ratio = survey_mean(afford_ratio, vartype = NULL, na.rm = TRUE)) %>%
  mutate(year = 2013) %>%
  ungroup()

summary2019 <- svy_afford_2019 %>%
  group_by(life_cycle_group) %>%
  summarise(mean_ratio = survey_mean(afford_ratio, vartype = NULL, na.rm = TRUE)) %>%
  mutate(year = 2019) %>%
  ungroup()

# ---- 3) Combine and coerce to numeric safely ----
# The survey_mean result creates a column named "mean_ratio" which should be numeric.
afford_summary <- bind_rows(summary2013, summary2019) %>%
# ensure numeric and remove any groups with NA mean
mutate(mean_ratio = as.numeric(mean_ratio)) %>%
  filter(!is.na(mean_ratio), life_cycle_group != "Not applicable")


# ---- 4) Plot as percentages, show value on top of bars, add 30% reference line ----
p <- ggplot(afford_summary, aes(x = life_cycle_group, y = mean_ratio, fill = factor(year))) +
  geom_col(position = position_dodge(width = 1.2), width = 0.5) +   # <- wider dodge, narrower bars
  geom_text(aes(label = paste0(round(mean_ratio * 100, 1), "%")),
            position = position_dodge(width = 1.2),
            vjust = -0.3, size = 3) +
  geom_hline(yintercept = 0.30, linetype = "dashed", color = "red") +
  annotate("text", x = Inf, y = 0.30, label = "30% stress threshold", 
           hjust = 1.05, vjust = -0.5, color = "red", size = 3) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, NA)) +
  labs(
    title = "Mean housing affordability ratio (weekly housing cost ÷ weekly income)\n2013 vs 2019 by life-cycle group",
    x = "Life-cycle group",
    y = "Mean housing cost as % of weekly income",
    fill = "Year"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1, size = 6),
    plot.title = element_text(size = 11, face = "bold")
  )

ggplotly(p)
```
```{r}
#Sought financial help from friends/family due to a shortage of money	CFFAMILY
# 0. Did not seek financial help from friends/family
# 1. Sought financial help from friends/family


#Ability of household to raise $500 emergency money  EMERGMON  0. Not applicable
# 1. Could raise $500 within a week
# 5. Could not raise $500 within a week
# 99. Not stated



# Extract underlying data frame
df <- as.data.frame(hh_svy$variables)

# Base data processing
fam_counts_only <- df %>%
  select(life_cycle_group, CFFAMILY, EMERGMON) %>%
  pivot_longer(cols = c(CFFAMILY, EMERGMON),
               names_to = "variable", values_to = "value") %>%
  # 💡 Filter: Keep only the CFFAMILY variable
  filter(variable == "CFFAMILY", value == 1, life_cycle_group != "Not applicable") %>%
  group_by(life_cycle_group, variable) %>%
  summarise(count_1 = n(), .groups = "drop") %>%
  # Rename variable to neat label
  mutate(variable = dplyr::recode(variable,
                           CFFAMILY = "Help from family")) %>%
  filter(count_1 > 0) %>%
  # Calculate percentages (relative to the total count for Help from family)
  group_by(variable) %>%
  mutate(perc = count_1 / sum(count_1) * 100,
         label = paste0(round(perc, 1), "%"))

em_counts_only <- df %>%
  select(life_cycle_group, CFFAMILY, EMERGMON) %>%
  pivot_longer(cols = c(CFFAMILY, EMERGMON),
               names_to = "variable", values_to = "value") %>%
  # 💡 Filter: Keep only the EMERGMON variable
  filter(variable == "EMERGMON", value == 1, life_cycle_group != "Not applicable") %>%
  group_by(life_cycle_group, variable) %>%
  summarise(count_1 = n(), .groups = "drop") %>%
  # Rename variable to neat label
  mutate(variable = dplyr::recode(variable,
                           EMERGMON = "Emergency money")) %>%
  filter(count_1 > 0) %>%
  # Calculate percentages (relative to the total count for Emergency money)
  group_by(variable) %>%
  mutate(perc = count_1 / sum(count_1) * 100,
         label = paste0(round(perc, 1), "%"))

# Data frame for 'Emergency money'
em_data <- em_counts_only %>%
  filter(variable == "Emergency money") %>%
  mutate(life_cycle_group = droplevels(as.factor(life_cycle_group)))

# Data frame for 'Help from family'
fam_data <- fam_counts_only  %>%
  filter(variable == "Help from family") %>%
  mutate(life_cycle_group = droplevels(as.factor(life_cycle_group)))

plot_native_pie <- function(data, plot_title) {
  # Ensure the percentages are used for the text display
  data <- data %>%
    mutate(life_cycle_group = as.factor(life_cycle_group))
  
  plot_ly(data, 
          labels = ~life_cycle_group, 
          values = ~count_1, 
          type = 'pie',
          # Set the text displayed on the slice, and the hover text
          text = ~label, 
          hovertemplate = paste('<b>%{label}</b><br>',
                                '%{value}<extra></extra>'),
          textinfo = 'text',
          insidetextfont = list(color = '#FFFFFF', size = 10),
          # RColorBrewer must be installed/loaded for this color line
          marker = list(line = list(color = '#FFFFFF', width = 1)),
          name = plot_title,
          domain = list(x = c(0, 1), y = c(0, 1))) %>% 
    layout(title = list(text = plot_title, font = list(size = 13)),
           showlegend = TRUE)
}

# --- GENERATE AND PRINT INTERACTIVE PLOTS ---

# 1. Generate the individual INTERACTIVE plots
# Note: Use plot_native_pie, NOT plot_pie
plot_em_interactive <- plot_native_pie(em_data, "No enough Emergency money")
plot_fam_interactive <- plot_native_pie(fam_data, "Need financial Help from family")

# 2. Print the Emergency Money interactive plot
print(plot_em_interactive)

# 3. Print the Help from Family interactive plot
print(plot_fam_interactive)
```


```{r}
#Weekly housing costs (broad) as a percentage of gross hh income (exc RA) ranges: weekly_housing_cost_cat
#Number of persons in household:	PERSHBC
#Level of highest educational attainment: LVLEDUA
#Tenure type - HH: TENURECF
#Age of youngest dependent child in household:	AGEKIDH
#State or Territory of usual residence 2016: STATEHEC
#Purchase price of dwelling (for dwellings purchased/built in the previous 3 years):HVALPRIC


explore2019<- svy_afford_2019 %>%
  select(
    weekly_housing_cost    = HCOSTSH,
    weekly_income          = hh_income_from_ius,
    HousingCostRatio_Range = weekly_housing_cost_cat,
    Life_Cycle_Group       = life_cycle_group, 
    Household_Size         = PERSHBC,
    Affordability_Ratio    = afford_ratio,
    Education_Level        = LVLEDUA,
    Age_Youngest_Child     = AGEKIDH,
    State                  = STATEHEC,
    Tenure_type            = TENURECF,
    Purchase_price_dwelling= HVALPRIC
  ) %>%
   # Filter out Tenure type == 0 (Not applicable)
  filter(Tenure_type != 0)  %>%
  # Replace 999 in Age_Youngest_Child with mean of valid ages (0–24)
  mutate(Age_Youngest_Child = ifelse(
      Age_Youngest_Child == 999,
      mean(Age_Youngest_Child[Age_Youngest_Child != 999], na.rm = TRUE),
      Age_Youngest_Child
    ),
    #convert into factor vars
    State = as.factor(State),
    Tenure_type = as.factor(Tenure_type),
    Education_Level = as.factor(Education_Level),
    # Recode Weekly housing cost ratio variable:
    # 0 if <3 (less than 30% of income), 1 if >=3 (more than 30%)
    HousingStress = ifelse(HousingCostRatio_Range < 3, 0, 1))
    
  
  
```



```{r}


model <- glm(HousingStress ~ Life_Cycle_Group + Household_Size + 
                     Education_Level + Age_Youngest_Child + State + Tenure_type + 
                     Purchase_price_dwelling + weekly_housing_cost + weekly_income,  
                       data = explore2019,
                       family = binomial(link = "logit")) %>% 
  step(trace=FALSE)
summary(model)
car::vif(model)
```

```{r}
model2 <- glm(HousingStress ~ Life_Cycle_Group + Household_Size + 
                     Education_Level + Age_Youngest_Child + State + Tenure_type + 
                     Purchase_price_dwelling + weekly_housing_cost * weekly_income,  
                       data = explore2019,
                       family = binomial(link = "logit")) %>% 
  step(trace=FALSE)
summary(model2)
car::vif(model2)
```
The main predictors are:

weekly_housing_cost ↑ → higher stress

weekly_income ↑ → lower stress

weekly_housing_cost × weekly_income → the effect of cost depends on income level (interaction)

Tenure_type and Purchase_price_dwelling also matter.

Interaction interpretation:

For low-income households, rising housing costs rapidly increase stress.

For high-income households, that effect is smaller — they can absorb higher costs.

```{r}
#deviance test
anova(model, test="Chisq")
drop1(model, test="Chisq")

anova(model2, test="Chisq")
drop1(model2, test="Chisq")

anova(model, model2, test="Chisq")
```


```{r}
AIC(model, model2)
```

```{r}
#Diagnostics 

aug <- augment(model2)
ggplot(aug, aes(x=.fitted, y=.resid)) +
geom_point() + geom_smooth()

ggplot(aug, aes(sample=.resid)) +
 geom_qq()
```

