---
title: |
  <br>**Housing Stress in Australia: Trends, Drivers, and Insights**
author: |
  | Author: Chenghao Zeng
  | Supervisor: Rebecca Valenzuela, Australian Institute of Family Studies
date: "25/10/2025"
output: rmdformats::readthedown
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo=FALSE, 
                      message=FALSE,
                      error = FALSE,
                      warning=FALSE,
                      eval= TRUE,
                      fig.align = "center")
```

```{r include=FALSE}

library(tidyverse)
library(survey)
library(srvyr)
library(scales)
library(glue)
library(plotly)
library(patchwork)
library(nnet)
library(dplyr)
library(tidyr)
library(car)
library(brglm2)
library(logistf)
library(visreg)
library(broom)
library(mice)
```

# Git Repo Link: https://github.com/ChenghaoZeng/5543project/tree/main

# Abstract

This report examines the evolving patterns of housing stress in Australia between 2013 and 2019, identifying the most vulnerable demographic groups and exploring the key factors that drive financial strain related to housing costs. Using national household survey data, the study applies both descriptive and inferential analysis — including logistic regression — to reveal structural inequalities that contribute to housing stress. Key findings indicate that lone-person households under 35, single-parent families, and female-headed households remain consistently at higher risk. The study also highlights how factors such as tenure type, purchase price, and household size significantly influence the likelihood of housing stress. The insights generated aim to inform policymakers and contribute to discussions on targeted housing affordability interventions.


# Introduction

Housing affordability has become one of the most pressing social and economic challenges in Australia. The growing mismatch between housing supply and population growth has placed considerable pressure on both renters and prospective homeowners. The concept of housing stress — commonly defined as households spending more than 30% of their gross income on housing costs — has become a widely used indicator of financial vulnerability (ABC News, Hutchens, 2024).

The purpose of this report is to explore how housing stress has evolved over time, to identify which groups are most affected, and to uncover the underlying drivers of this financial burden. The analysis also aims to create accessible and interpretable visual tools that can communicate findings effectively to policymakers and the public.

This work was conducted as part of an internship project at the Australian Institute of Family Studies, focusing on translating analytical evidence into policy-relevant insights.


# Background and Motivation : 

Australia’s housing landscape has experienced significant shifts over the past two decades. Escalating property prices, limited housing supply, and slow wage growth have created conditions where many households — particularly those with children or on lower incomes — struggle to meet housing costs.

According to the 2024 ABC News report, housing stress has “risen sharply since the early 2000s,” with more Australians spending disproportionate amounts of income on housing (Hutchens, 2024). This trend has profound implications for family wellbeing, social inclusion, and economic stability.

Beyond simple affordability concerns, housing stress affects family functioning, mental health, and intergenerational mobility. Families under persistent housing stress are more likely to face financial insecurity, relationship strain, and poorer health outcomes. These consequences make housing stress not merely an economic issue, but a broader social and policy challenge.

The motivation for this study arises from these concerns. By analysing national household survey data across multiple years, this project aims to clarify which groups face the highest burden and why — thereby helping to inform future housing and welfare policy.

# Objectives and Significance

The analysis has four main objectives:

1: To describe how housing stress levels have changed over time (2013–2019).

2: To identify which demographic and household groups are most at risk.

3: To model key drivers — such as household structure, education, tenure type, and financial variables — that explain housing stress.

4: To translate findings into practical, policy-relevant insights.

This study was undertaken as part of my internship at the Australian Institute of Family Studies (AIFS), an independent statutory agency of the Australian Government within the Department of Social Services. AIFS conducts high-quality research to increase understanding of the factors that affect family wellbeing and to inform government policy and service delivery. Its work covers diverse areas such as family relationships, housing and financial security, child wellbeing, and social inclusion.

As an intern analyst at AIFS, my role focused on using data-driven methods to explore issues affecting Australian families, particularly in relation to housing affordability and economic stress. This project was designed not only as an academic exercise but as a practical contribution to AIFS’s mission of supporting evidence-based policy.

The analysis contributes to this broader goal by producing reproducible, interpretable, and visually engaging insights that can be shared with policymakers, researchers, and the public.

The significance of this study lies in its combination of statistical modelling and visual storytelling. By blending quantitative rigour with clear, visual communication, the analysis aims to:

+ Enhance understanding of complex affordability dynamics among different population groups,

+ Inform evidence-based decision making on housing and social welfare policy, and

+ Provide an analytical framework that can be extended with future data — for example, to assess housing stress patterns in the post-COVID period.

Ultimately, this project bridges the gap between academic research and policy application, demonstrating how analytical insights can be translated into meaningful recommendations that improve family and community wellbeing across Australia.

# Methodology

## Data Processing

### Overview

The Statistical Income and Housing (SIH) microdata for 2013 and 2019 were prepared and harmonised into household-level analysis files. The following sections document the steps performed to construct the analysis datasets, the rationale behind key transformations, and important implementation checks. All operations are described in a neutral, reproducible manner and correspond to the R code used in the analysis pipeline.


### 1. Loading raw files
Household-level, person-level, income and loan CSV files were read for both survey years. The files correspond to the SIH file components:

+ BH — household-level file

+ BP — person-level file

+ BI — income (income unit / person) file

+ BL — loans / borrowings file

```{r, eval=FALSE}
#load data
household <- read_csv("data/CSV/SIH19BH.csv", show_col_types = FALSE)
income <- read_csv("data/CSV/SIH19BI.csv", show_col_types = FALSE)
loans <- read_csv("data/CSV/SIH19BL.csv", show_col_types = FALSE)
person <- read_csv("data/CSV/SIH19BP.csv", show_col_types = FALSE)

#load older data
household2013 <- read_csv("data/CSV/SIH13BH.csv", show_col_types = FALSE)
income2013 <- read_csv("data/CSV/SIH13BI.csv", show_col_types = FALSE)
loans2013 <- read_csv("data/CSV/SIH13BL.csv", show_col_types = FALSE)
person2013 <- read_csv("data/CSV/SIH13BP.csv", show_col_types = FALSE)
```

+ Reading both 2013 and 2019 prepares the data for direct comparison across years.

+ read_csv(..., show_col_types = FALSE) suppresses the printing of column type messages for cleaner logs while preserving column types.

### 2. Aggregating loan records to household level

Loan records are at the loan-level and were aggregated to household totals using the household identifier (ABSHID). The aggregated measures produced are:

+ hh_total_weekly_repayments — sum of weekly repayments per household

+ hh_total_outstanding — sum of outstanding loan balances per household

+ hh_num_loans — number of loan records associated with the household


```{r, eval=FALSE}
#Aggregate loans (household-level debt measures)

loans_agg <- loans %>%
  group_by(ABSHID) %>%
  summarise(
    hh_total_weekly_repayments = sum(WKREPLN, na.rm = TRUE),
    hh_total_outstanding = sum(AMTOLN, na.rm = TRUE),
    hh_num_loans = n(),
    .groups = "drop"
  )


loans_agg2013 <- loans2013 %>%
  group_by(ABSHID) %>%
  summarise(
    hh_total_weekly_repayments = sum(WKREPLN, na.rm = TRUE),
    hh_total_outstanding = sum(AMTOLN, na.rm = TRUE),
    hh_num_loans = n(),
    .groups = "drop"
  )
```

+ Aggregation converts loan-level detail into household-level covariates suitable for modelling and descriptive statistics.

+ na.rm = TRUE prevents missing loan entries from producing NA sums; the semantic meaning of missing values should be assessed (e.g., missing vs no loans).

### 3. Aggregating income-unit / income records to household level

Income records were aggregated by household to create an auxiliary household-level income measure and a count of income units:
```{r, eval=FALSE}
iu_agg <- income %>%
group_by(ABSHID) %>%
summarise(
hh_income_from_ius = sum(INCTSCU8, na.rm = TRUE),
hh_num_ius = n(),
.groups = "drop"
)

iu_agg2013 <- income2013 %>%
group_by(ABSHID) %>%
summarise(
hh_income_from_ius = sum(INCTSCU8, na.rm = TRUE),
hh_num_ius = n(),
.groups = "drop"
)
```

+ Summed income from income units provides an independent check against the household income variable in the household file and supports alternative income constructions when required.

### 4. Extracting household reference person attributes

Person-level variables for the household reference person (HHPOS == 1) were selected and transformed into household-level indicators. A proxy for the health status of the household head was constructed as a binary flag:

```{r, eval=FALSE}
person_head <- person %>%
filter(HHPOS == 1) %>%
select(ABSHID, SAHQ01, AGEEC, SEXP, LVLEDUA) %>%
mutate(
head_poor_health_flag = case_when(
SAHQ01 %in% c(4,5) ~ 1L,
TRUE ~ 0L
)
)

person_head2013 <- person2013 %>%
filter(HHPOS == 1) %>%
select(ABSHID, DISBPER, AGEBC, SEXP) %>%
mutate(
head_poor_health_flag = case_when(
DISBPER == 2 ~ 1L,
TRUE ~ 0L
)
)
```

+ The household reference person provides demographic and wellbeing proxies that are informative at the household level.

+ Differences in variable names between years were handled explicitly to harmonise equivalent constructs.

### 5. Merging components into a household analysis file

Aggregated loan and income measures and the household reference person attributes were merged onto the household frame using left joins keyed by ABSHID:

```{r, eval=FALSE}
hh_merged <- household %>%
left_join(loans_agg, by = "ABSHID") %>%
left_join(iu_agg, by = "ABSHID") %>%
left_join(person_head, by = "ABSHID")

hh_merged2013 <- household2013 %>%
left_join(loans_agg2013, by = "ABSHID") %>%
left_join(iu_agg2013, by = "ABSHID") %>%
left_join(person_head2013, by = "ABSHID")
```

+ Left joins preserve the household sampling frame while attaching derived covariates. Households lacking particular records (e.g., no loans) appear with NA in the relevant fields, allowing for explicit treatment of missingness downstream.

### 6. Variable renaming and harmonisation

A mapping of original variable codes to descriptive names was applied to improve readability and reproducibility. Conditional renaming was used to accommodate differences in variable availability between years.
```{r, eval=FALSE}
rename_map <- c(
"SIHHHWT"   = "hh_weight",
"HCOSTH8A"  = "weekly_housing_cost_pct",
"HCOST10A"  = "weekly_housing_cost_cat",
"INCTSCH8"  = "hh_total_weekly_income",
"DISPSCH8"  = "hh_weekly_disposable_income",
"LIFECYCH"  = "life_cycle_group_code",
"DPKIDHBC"  = "num_dependents_under25",
"FSTRSS19"  = "num_financial_stress_events",
"EMGMONEY"  = "ability_raise_2000_flag"
)

for(orig in names(rename_map)){
if(orig %in% names(hh_merged)){
hh_merged <- hh_merged %>% rename(!!rename_map[[orig]] := all_of(orig))
}
}

for(orig in names(rename_map)){
if(orig %in% names(hh_merged2013)){
hh_merged2013 <- hh_merged2013 %>% rename(!!rename_map[[orig]] := all_of(orig))
}
}
```

+ Human-readable variable names facilitate interpretation of tables, figures and model specifications and reduce error when sharing code.

### 7. Recoding life-cycle groups and deriving outcome flags

Numeric life-cycle codes were recoded to descriptive labels. Multiple housing-stress indicators were constructed using different thresholds to support sensitivity analyses:

```{r, eval=FALSE}
hh_merged <- hh_merged %>%
mutate(
life_cycle_group = case_when(
life_cycle_group_code == 0  ~ "Not applicable",
life_cycle_group_code == 1  ~ "Lone person <35",
life_cycle_group_code == 2  ~ "Couple (ref <35)",
life_cycle_group_code == 3  ~ "Couple w dependents, eldest <5",
life_cycle_group_code == 4  ~ "Couple w dependents, eldest 5-14",
life_cycle_group_code == 5  ~ "Couple w dependents, eldest 15-24",
life_cycle_group_code == 6  ~ "One parent w dependents",
life_cycle_group_code == 7  ~ "Couple w dep & non-dep",
life_cycle_group_code == 8  ~ "Couple w non-dep only",
life_cycle_group_code == 9  ~ "Couple (ref 55-64)",
life_cycle_group_code == 10 ~ "Couple (ref 65+)",
life_cycle_group_code == 11 ~ "Lone person 65+",
TRUE ~ paste0("Other (", life_cycle_group_code, ")")
),
housing_stress_flag = case_when(
!is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct > 30 ~ 1L,
!is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct <= 30 ~ 0L,
TRUE ~ NA_integer_
),
housing_stress_flag_20 = case_when(
!is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct > 20 ~ 1L,
!is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct <= 20 ~ 0L,
TRUE ~ NA_integer_
),
housing_stress_flag_40 = case_when(
!is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct > 40 ~ 1L,
!is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct <= 40 ~ 0L,
TRUE ~ NA_integer_
),
housing_stress_flag_50 = case_when(
!is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct > 50 ~ 1L,
!is.na(weekly_housing_cost_pct) & weekly_housing_cost_pct <= 50 ~ 0L,
TRUE ~ NA_integer_
),
ability_raise_2000 = case_when(
ability_raise_2000_flag == 1 ~ "Can raise $2000",
ability_raise_2000_flag == 5 ~ "Cannot raise $2000",
TRUE ~ NA_character_
)
)
```

+ Descriptive life-cycle labels reduce reliance on codebooks in figures and tables.

+ Multiple housing-stress flags enable sensitivity analysis to threshold choice and facilitate characterization of varying degrees of cost burden.


### 8. Constructing survey replicate design objects

Replicate-weight columns were identified programmatically and survey-replicate designs were created using srvyr::as_survey_rep with Fay’s method (rho = 0.5):

```{r,  eval=FALSE}
rep_cols <- grep("^WHS\d{2}$", names(hh_merged), value = TRUE)

hh_svy <- hh_merged %>%
as_survey_rep(weights = hh_weight,
repweights = all_of(rep_cols),
type = "Fay",
rho = 0.5)
```
An equivalent procedure was used for the 2013 dataset, with a year-specific pattern for the replicate-weight column names.

Rationale and benefits:

+ The SIH is a complex survey; proper variance estimation requires replicate weights rather than naive standard errors from simple random-sample assumptions.

+ Replicate weights supplied by the ABS reflect the original complex design and permit design-consistent variance estimation.

+ Fay’s method is a BRR variant that reduces perturbation extremes and tends to produce more stable variance estimates; rho = 0.5 is a conventional Fay factor for ABS replicate weights.

+ The srvyr interface integrates survey-package variance estimation with tidyverse-style pipelines, enabling readable and reproducible analysis code (e.g., group_by() and summarise() combined with survey_mean() and survey_total()).


# Exploratory Data Analysis

### Minimum legal age of marriage without consent
**Aim:** To explore countries where the minimum legal age of marriage is less than 18

**About:** This dataset is part of the Minimum Gender Dataset compiled by the United Nations Statistics Division.

```{r}



```

**Observations:** 
We can observe that most countries where the legal marriage age is even less than 18 are mostly Asian countries and European island countries. It is mostly observed that Asian countries have long-term traditions and customs which include social stigmas like child marriage due to which such issues still exist in society. Many countries which are dominantly Islamic have high rates of child marriages due to customs existing within religion. 

**Source of Dataset:** [United Nations](https://gender-data-hub-2-undesa.hub.arcgis.com/datasets/8ba8255a4ba047fcb0e6dbe8041a0eb0/explore)

### Covid-19 Stringency

**About:** The stringency index is a composite measure based on nine response indicators including school closures, workplace closures, and travel bans, rescaled to a value from 0 to 100 (100 = strictest). Timeline: 2020-2022

**Aim:** To explore which countries had the strictest government policies in terms of school closures and other lockdown measures

```{r}



```

**Observations:** 
We can observe most South Asian countries and North African countries has high stringency index. 
China has the highest stringency index. 
In contrast, it is observed most Central African countries like Chad, Mali, and Asian countries like Yemen and Afghanistan had the lowest stringency index, and yet these countries have the highest number of child marriages globally.  

**Source of Dataset:** [Our World in Data](https://ourworldindata.org/metrics-explained-covid19-stringency-index)



### Legal Policies Regarding Child Marriage 

**Aim:** To understand the policies regarding child marriage globally. Timeline: 2019

**Data Dictionary:** spec_tbl_df [131 x 33]

* iso_a3  : ISO3 Country Code
* Entity  : Country Name
* Year    : Year
* Law     : Legal code

```{r warning=FALSE}

```

**Observations:**

+ Even most advanced countries like the USA allow child marriage.
+ Nordic countries like Switzerland, Denmark, and Sweden are the only countries that completely prohibit child marriage irrespective of any customs or legal exceptions.
+ Middle-East countries which are mostly Islamic dominant countries like Yemen, Saudi Arabia, and Iran, and African countries like Sudan allow child marriages.
+ Most African countries like Mali, and Tanzania and, Asian countries like Taiwan, Malaysia, Pakistan, and Afghanistan allow girl-child marriages.

**Source:** [Our World in Data](https://ourworldindata.org/grapher/laws-on-child-marriage)

### Global Child Marriage Rate

**About:** This dataset is part of the Global SDG Indicator Database compiled through the UN System in preparation for the Secretary-General's annual report on Progress towards Sustainable Development Goals. This dataset contains the latest value of the child marriage rate from 2005-2022.
Indicator 5.3.1: Proportion of women aged 20–24 years who were married or in a union before age 15 and before age 18
Target 5.3: Eliminate female social issues like child marriage, female gender mutilation
Goal 5: Achieve gender equality and empower all women and girls

**Data Dictionary:** spec_tbl_df [131 x 33]
* geoAreaCode : num [1:131] Country Code
* geoAreaName : chr [1:131] Country Name
* level_      : num [1:131] Numeric
* parentCode  : num [1:131] Numeric
* parentName  : chr [1:131] Sub Continental Region
* type        : chr [1:131] Geographical level
* X           : num [1:131] Latitude
* Y           : num [1:131] Longitude
* ISO3        : chr [1:131] ISO3 country code
* latest_value: num [1:131] Child Marriage Rate

```{r data}


```


```{r visualisation}



```


**Observations:**

+ We can observe most African countries have some of the highest rates of child marriages with countries like Chad, and Niger displaying the highest rate followed by Sudan, Mozambique, Nigeria, etc.
+ Among Asian countries, Bangladesh has the highest child marriage rate.

**Source:** [UN DESA Statistics Division](https://unstats-undesa.opendata.arcgis.com/datasets/undesa::indicator-5-3-1-proportion-of-women-aged-20-24-years-who-were-married-or-in-a-union-before-age-18-percent/explore?location=2.960853%2C1.735981%2C2.67)


### Child Marriage Rate in Sub-continental region during Covid-19

**About:** This dataset contains the Percentage of women (aged 20-24 years) married or in union before age 18 during the Covid-19 period in sub-continental regions. 

```{r echo=FALSE, warning=FALSE, message= FALSE}


```


**Observation:** We can observe that African countries show the highest rate of child marriage during Covid-19 followed by South Asian countries and Latin American countries. 

**Data Source:** [UNICEF Data](https://data.unicef.org/resources/data_explorer/unicef_f/?ag=UNICEF&df=GLOBAL_DATAFLOW&ver=1.0&dq=.PT_F_20-24_MRD_U15+PT_F_20-24_MRD_U18+PT_F_15-19_MRD..&startPeriod=2016&endPeriod=2022)


## Final Visualisation: % women aged 20-24 married before 18 (average 2005-2020) and the legal policy regarding child marriage


```{r}

```

### 




# Discussion
5.1 Interpreting the Patterns

The findings paint a nuanced picture of Australia’s housing stress landscape. Although some improvements occurred between 2013 and 2019, the benefits were unevenly distributed.

+ Younger and single-parent households remained the most financially vulnerable, reflecting unstable income sources and limited access to affordable housing.

+ Gender inequality persisted, showing that broader economic disparities spill over into housing outcomes.

+ Household size and composition played a protective role — highlighting the importance of shared financial support systems.

These patterns suggest that housing stress is both a financial and social phenomenon, embedded in broader structures of inequality.

# Limitations and Future Work

The main limitation of this analysis is the data coverage period, which ends in 2019 — prior to the COVID-19 pandemic. Housing markets and affordability patterns have changed substantially since then, influenced by rising interest rates, inflation, and shifts in remote work.

Future extensions will include:

+ Updating the analysis with post-2020 data to assess pandemic-era impacts;

+ Integrating regional housing market data to explore spatial inequalities; and

+ Considering dynamic models (e.g., mixed effects) to account for household-level variation over time.

This continuation will enable a more complete understanding of housing stress in contemporary Australia.

## Conclusion 

This report provides an evidence-based assessment of housing stress trends in Australia, identifying persistent disparities by age, family structure, and gender.
Despite modest improvements between 2013 and 2019, key groups — notably lone young adults, single parents, and female-headed households — continue to face disproportionate financial burdens.

The modelling results demonstrate that tenure type, household composition, and dwelling costs are among the most critical predictors of stress. These insights underscore the need for targeted housing policies and financial assistance mechanisms that account for structural inequalities.

Ultimately, addressing housing stress requires a multi-pronged policy approach — one that not only expands affordable supply but also improves financial resilience among vulnerable households.




# References

+ UNICEF. (2021) 10 million additional girls at risk of child marriage due to COVID-19, UNICEF. Available at: https://www.unicef.org/eap/press-releases/10-million-additional-girls-risk-child-marriage-due-covid-19-unicef

+ Onabanjo, J., Kalasa, B., & Abdel-Ahad, M. (2014, May 28). OP-ED: ENDING CHILD MARRIAGE IN AFRICA CAN NO LONGER WAIT. Global Information Network Retrieved from https://www.proquest.com/wire-feeds/op-ed-ending-child-marriage-africa-can-no-longer/docview/1534961300/se-2

+ KAMAL, S. M. M., HASSAN, C. H., ALAM, G. M., & YING, Y. (2015). CHILD MARRIAGE IN BANGLADESH: TRENDS AND DETERMINANTS. Journal of Biosocial Science, 47(1), 120-39. doi:https://doi.org/10.1017/S0021932013000746

+ Afrin, Tangina, & Zainuddin, Mohammad. (2021). Spike in child marriage in Bangladesh during COVID-19: Determinants and interventions. Child Abuse & Neglect, 112, 104918–104918. https://doi.org/10.1016/j.chiabu.2020.104918

+ The economy of wellbeing: what is it and what are the implications for health? BMJ 2020; 369 doi: https://doi.org/10.1136/bmj.m1874

+ Pebesma, E., & Bivand, R. (2023). Spatial Data Science: With Applications in R (1st
  ed.). Chapman and Hall/CRC. https://doi.org/10.1201/9780429459016

+ Zeileis A, Fisher JC, Hornik K, Ihaka R, McWhite CD, Murrell P, Stauffer R, Wilke CO
(2020). “colorspace: A Toolbox for Manipulating and Assessing Colors and Palettes.”
_Journal of Statistical Software_, *96*(1), 1-49. doi: 10.18637/jss.v096.i01 (URL:
https://doi.org/10.18637/jss.v096.i01).

+ Chan, C., Geoffrey CH Chan, Leeper, T.J. and Becker, J. (2021). rio: A Swiss-army knife for data file I/O.
+ FC, M., Davis, T.L. and ggplot2 authors (2022). ggpattern: ‘ggplot2’ pattern geoms. [online] Available at: https://cran.r-project.org/package=ggpattern.
+ Firke, S. (2023). janitor: Simple tools for examining and cleaning dirty data. [online] Available at: https://cran.r-project.org/package=janitor.
+ Garnier, Simon, Ross, Noam, Rudis, Robert, Camargo, Pedro, A., Sciaini, Marco, Scherer and Cédric (2021). viridis - colorblind-friendly color maps for r. [online] doi:https://doi.org/10.5281/zenodo.4679424.
+ Massicotte, P. and South, A. (2023). rnaturalearth: World map data from natural earth. [online] Available at: https://cran.r-project.org/package=rnaturalearth.
Ooms, J. (2022). gifski: Highest quality GIF encoder. [online] Available at: https://cran.r-project.org/package=gifski.
+ Sievert, C. (2020). Interactive web-based data visualization with r, plotly, and shiny. [online] Chapman and Hall/CRC. Available at: https://plotly-r.com.
+ Tierney, N. and Cook, D. (2023). Expanding tidy data principles to facilitate missing data exploration, visualization and assessment of imputations. Journal of Statistical Software, 105, pp.1–31. doi:https://doi.org/10.18637/jss.v105.i07.
+ Wickham, H. (2016). ggplot2: Elegant graphics for data analysis. [online] Springer-Verlag New York. Available at: https://ggplot2.tidyverse.org.
+ Wickham, H., Averick, M., Bryan, J., Chang, W., Lucy D'Agostino McGowan, Romain François, Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Thomas Lin Pedersen, Miller, E., Stephan Milton Bache, Kirill Müller, Ooms, J., Robinson, D., Dana Paige Seidel, Vitalie Spinu and Takahashi, K. (2019). Welcome to the tidyverse. Journal of Open Source Software, 4, p.1686. doi:https://doi.org/10.21105/joss.01686.
+ Wickham, H. and Bryan, J. (2023). readxl: Read excel files. [online] Available at: https://cran.r-project.org/package=readxl.
+ Wickham, H., Hester, J. and Bryan, J. (2023). readr: Read rectangular text data. [online] Available at: https://cran.r-project.org/package=readr.

